Weakly Nonlinear MRI: Notes

### 2014-03-04 notes/mri_notes.tex
fast and slow vars

Linearizing our four equations, we find a system described by the matrix: \\

%$\left[\begin{smallmatrix}{}s \left(- k_{x}^{2} - k_{z}^{2}\right) - \frac{1}{Re} \left(k_{x}^{4} + 2 k_{x}^{2} k_{z}^{2} + k_{z}^{4}\right) & 2 i k_{z} & \frac{i B_{0} k_{z}}{4 \pi} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0\\\i k_{z} \\left(- \\Omega_{0} q + 2\\right) & s - \\frac{1}{Re} \\left(- k_{x}^{2} - k_{z}^{2}\\right) & 0 & - \\frac{i B_{0} k_{z}}{4 \\pi}\\\\- i B_{0} k_{z} & 0 & s - \\frac{1}{Rm} \\left(- k_{x}^{2} - k_{z}^{2}\\right) & 0\\\\0 & - i B_{0} k_{z} & i \\Omega_{0} k_{z} q & s - \\frac{1}{Rm} \\left(- k_{x}^{2} - k_{z}^{2}\\right)\end{smallmatrix}\right]$

\begin{equation*}
\left[\begin{matrix}
\left(- k_{x}^{2} - k_{z}^{2}\right) - \frac{1}{Re} \left(k_{x}^{4} + 2 k_{x}^{2} k_{z}^{2} + k_{z}^{4}\right) & 2 i k_{z} & \frac{i B_{0} k_{z}}{4 \pi} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0\\

@@

\end{matrix}\right] \cdot \left[\begin{matrix}

@@

>>>>

Linearizing our four equations, we find a system described by the matrix: \\
\begin{equation*}
M = \left[\begin{matrix}
\left(- k_{x}^{2} - k_{z}^{2}\right) - \frac{1}{Re} \left(k_{x}^{4} + 2 k_{x}^{2} k_{z}^{2} + k_{z}^{4}\right) & 2 i k_{z} & \frac{i B_{0} k_{z}}{4 \pi} \left(- k_{x}^{2} - k_{z}^{2}\right) & 0\\

@@

where: 
$M \cdot \left[\begin{matrix}

@@

\end{matrix}\right] = \left[\begin{matrix}
0 \\
0\\
0\\
0\\
\end{matrix}\right]
$ \\

The determinant of M gives the dispersion relation: \\

$\frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q s^{2} - \frac{4 s}{Rm} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 s}{Rm} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{Rm^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{Rm^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{Rm^{2}} k_{z}^{6} q - k_{x}^{2} s^{4} - k_{z}^{2} s^{4} + 4 k_{z}^{2} s^{2} - \frac{2 s^{3}}{Rm} k_{x}^{4} - \frac{4 s^{3}}{Rm} k_{x}^{2} k_{z}^{2} + \frac{8 s}{Rm} k_{x}^{2} k_{z}^{2} - \frac{2 s^{3}}{Rm} k_{z}^{4} + \frac{8 s}{Rm} k_{z}^{4} - \frac{k_{x}^{6} s^{2}}{Rm^{2}} - \frac{3 s^{2}}{Rm^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{Rm^{2}} k_{z}^{2} - \frac{3 s^{2}}{Rm^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{Rm^{2}} k_{z}^{4} - \frac{k_{z}^{6} s^{2}}{Rm^{2}} + \frac{4 k_{z}^{6}}{Rm^{2}} - \frac{2 s^{3}}{Re} k_{x}^{4} - \frac{4 s^{3}}{Re} k_{x}^{2} k_{z}^{2} - \frac{2 s^{3}}{Re} k_{z}^{4} - \frac{4 k_{x}^{6} s^{2}}{Re Rm} - \frac{12 k_{x}^{4} k_{z}^{2} s^{2}}{Re Rm} - \frac{12 k_{x}^{2} k_{z}^{4} s^{2}}{Re Rm} - \frac{4 k_{z}^{6} s^{2}}{Re Rm} - \frac{2 k_{x}^{8} s}{Re Rm^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} s}{Re Rm^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} s}{Re Rm^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} s}{Re Rm^{2}} - \frac{2 k_{z}^{8} s}{Re Rm^{2}} - \frac{k_{x}^{6} s^{2}}{Re^{2}} - \frac{3 s^{2}}{Re^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 s^{2}}{Re^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} s^{2}}{Re^{2}} - \frac{2 k_{x}^{8} s}{Re^{2} Rm} - \frac{8 k_{x}^{6} k_{z}^{2} s}{Re^{2} Rm} - \frac{12 k_{x}^{4} k_{z}^{4} s}{Re^{2} Rm} - \frac{8 k_{x}^{2} k_{z}^{6} s}{Re^{2} Rm} - \frac{2 k_{z}^{8} s}{Re^{2} Rm} - \frac{k_{x}^{10}}{Re^{2} Rm^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{Re^{2} Rm^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{Re^{2} Rm^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{Re^{2} Rm^{2}} - \frac{5 k_{x}^{2} k_{z}^{8}}{Re^{2} Rm^{2}} - \frac{k_{z}^{10}}{Re^{2} Rm^{2}}$ \\

Which, as expected, is fourth order in s. We see that the spatial eigenvalues appear as $k_x^2$ and $k_z^2$ at lowest order. The temporal eigenvalue appears as $s$ at lowest order. We choose the scalings between fast variables $(x, y, z, t)$ and slow variables $(X, Y, Z, T)$ such that all eigenvalues come in at $\mathcal{O}(\epsilon^2)$. Hence we define: \\

\noindent $X = \epsilon x$, \, \, $Y = \epsilon y$, \, \, $Z = \epsilon z$, \, \, $T = \epsilon^2 t$ \\
